\chapter*{Acknowledgments}
\pdfbookmark{Acknowledgments}{acknowledgments} % Sets a PDF bookmark for the acknowledgment
\label{chap:acknowledgments}

The journey towards a PhD degree is never smooth.
To all the people who have appeared in my life during my six years of study, I would like to express my gratitude.
Thanks for joys, sorrows and all my experiences at Columbia University that enabled me to grow, to know myself and to explore the meaning of life.

On top of my gratitude lists are my supervisors: Prof. Andrei Beloborodov and Prof Lam Hui.
I am indebted to their great instructions in research as well as wise personal and career advices. 
If I need to choose one piece of advice from which I benefited most, it will be the high standard of research: always aiming at the highest scientific goal of advancing human knowledge.
As Andrei always tells me, ``when you write a paper, think of what a reader in fifty or a hundred years can learn from it.''

I would like to thank the faculty members and postdocs whom I have collaborated with in the department of physics and the department of astronomy.
Gratitudes to Prof. Yuri Levin, Prof. Greg Bryan, Prof Brian Metzger and Prof. Lorenzo Sironi for knowledge in physics and astronomy as well as help for my job applications and special gratitude to Prof. Mal Ruderman for long chats of wisdom.
Thanks to Jonathan Zrake, Luca Comisso, Daniel Siegel, Sam Wong and Luca Santoni for inspiring discussions at the theory center.

I would also express my sincere gratitude to students in the Pupin Hall with whom we study, discuss and have fun together.
Thanks to Alex Chen and Xiao Xiao for guidance and help in the graduate study.
Thanks to Ben Margalit, Ashley Bransgrove and Dhruv Desai for working together in the field of high energy astrophysics.
Special thanks to the following friends: Rui Hu, Yun Zhang, Zhenghan Gao, Tianhao Ren, Minghao Cheng, Chih-Hsi Lee and Zimo Sun, with whom we spent many days and nights playing cards at the graduate student lounge.

Finally I would like to thank my parents. Thank you for support and forgiveness of my inability to frequently visit home to fulfill my duty of filial piety. This thesis is dedicated to you.

\newpage \vspace*{8cm}
\pdfbookmark{Dedication}{dedication} % Sets a PDF bookmark for the dedication
\begin{center}
\large To the my parents %, Hua Gu and Jianhua Li
\end{center}
% Local Variables:
% TeX-master: "../thesis"
% End:
